---
title: "Homework 3"
name: "Chloe Chah"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include = FALSE}

library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(patchwork)
library(ggplot2)
library(hexbin)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)

```


### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by users. There are user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

#### How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

It seems like the top 3 aisles where the most items are ordered from are "fresh vegetables", "fresh fruits", and "packaged vegetable fruits". 

#### MMake a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 
```

#### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```


#### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers. 

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```

### Problem 2 

#### Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_df = 
  read_csv("./Data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "minutes", 
    names_prefix = "activity_", 
    values_to = "activity_count"
  ) %>% 
  mutate(minutes = as.numeric(minutes), 
         day_num = recode(day, "Monday" = "1", 
                             "Tuesday" = "2", 
                             "Wednesday" = "3", 
                             "Thursday" = "4", 
                             "Friday" = "5", 
                             "Saturday" = "6", 
                             "Sunday" = "7")
  ) %>%
  mutate(
    weekday_vs_weekend = case_when(
      day_num <=5 ~ "Weekday", 
      day_num > 5 ~ "Weekend", 
      TRUE ~ ""
    )
  ) %>%
  
  select(-day_num)

head(accel_df)

```

The accelarator dataset contains information regarding the following variables: `r ls(accel_df)`. There are a total of `r nrow(accel_df)` rows and `r ncol(accel_df)` columns in the final dataset. 

#### Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent? 

```{r}
accel_df %>%
  group_by(week, day_id) %>%
  summarize(
    total_activity = sum(activity_count)
  ) %>%
  knitr::kable()
```

The patient seems to be most active on Fridays and the least active on Saturdays. 

#### Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph. 

```{r}
accel_df %>%
ggplot(aes(x = minutes, y = activity_count, color = day)) +
  geom_line()
```

Based on the graph, it seems like the patient is more active during the day on Sundays and more active in the evening on Fridays. A noticeable trend is that the time at which the patient is more active seems to depend on which day of the week it is. 

### Problem 3

#### Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

```{r}

data(na_noaa)

ny_noaa_df = ny_noaa %>%
          separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
        mutate(
          year = factor(year), 
          month = as.integer(month), 
          day = as.integer(day),
          prcp = as.numeric(prcp),
          tmax = as.numeric(tmax), 
          tmin = as.numeric(tmin), 
          prcp = prcp/10, 
          tmax = tmax/10, 
          tmin = tmin/10
        ) 

head(ny_noaa_df)

```

#### For snowfall, what are the most commonly observed values? Why?

```{r}

snow_df = ny_noaa_df %>%
          count(snow) %>%
          arrange(desc(n))

head(snow_df)

```

It seems like the most commonly observed values was 0 when we count the observations of the snow variable and arrange it in a descending order. 

#### Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}

ny_noaa_df %>%
  filter(month %in% c("1", "7")) %>%
  group_by(id, year, month) %>%
  summarize(
    avg_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = avg_tmax, group = id, color = month))+
  geom_point(alpha = .5, size = .5) + 
  geom_path() + 
  facet_grid(.~month) +
  theme(axis.text.x = element_text(angle = 270, vjust = .5, hjust =1))

head(ny_noaa_df)

```

It seems like the average maximum temperature is generally much higher in July than in January over all the weather stations, which makes sense as they are in different seasons. There seems to be outliers in January of 1982 and July of 1988. 

#### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option). (ii) Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.; 

```{r}

tmax_tmin = ny_noaa_df %>%
            ggplot(aes(x = as.numeric(tmax), y = as.numeric(tmin))) +
            geom_hex() + 
            labs(
              x = "Maximum temperature (C)", 
              y = "Minimum temperature (C)"
            )+ 
            theme(legend.position = "none")

snow_dist = ny_noaa_df %>%
            filter(snow > 0 & snow <100) %>%
            ggplot(aes(x = year, y = snow, color = year)) + 
            geom_violin(alpha = .5) + 
            theme(axis.text.x = element_text(angle = 270, vjust = .5, hjust = 1)) + 
            theme(legend.position = "none") + 
            labs(
              x = "Year", 
              y = "Snowfall (mm)"
            )

tmax_tmin + snow_dist 

```



