---
title: "Homework 3"
output: github_document
---

```{r setup, include = FALSE}

library(tidyverse)
library(p8105.datasets)
library(ggridges)

```

### Problem 3

#### Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

```{r}

data(na_noaa)

ny_noaa_df = ny_noaa %>%
          separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
        mutate(
          year = factor(year), 
          month = as.integer(month), 
          day = as.integer(day),
          prcp = as.integer(prcp),
          tmax = as.integer(tmax), 
          tmin = as.integer(tmin), 
          prcp = prcp/10, 
          tmax = tmax/10, 
          tmin = tmin/10
        ) 

```

#### For snowfall, what are the most commonly observed values? Why?

```{r}

snow_df = ny_noaa_df %>%
          count(snow) %>%
          arrange(desc(n))

```

It seems like the most commonly observed values was 0 when we count the observations of the snow variable and arrange it in a descending order. 

#### Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}

ny_noaa_df %>%
  filter(month %in% c("1", "7")) %>%
  group_by(id, year, month) %>%
  summarize(
    avg_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = avg_tmax, group = id))+
  geom_point(alpha = .5, size = .5) + 
  geom_path() + 
  facet_grid(.~month) +
  theme(axis.text.x = element_text(angle = 270, vjust = .5, hjust =1))

```

It seems like the average maximum temperature is generally much higher in July than in January, which makes sense as they are in different seasons. 

#### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); 

```{r}



```

#### (ii) Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}


```

